---
title: "R Notebook"
output: html_notebook
---

# Group 10

*Libraries needed* 


# Biological question and Project Scope

# Clustering

```{r setup, include=FALSE}
library(dplyr);
library("stringr");
library(ggplot2);
library(dplyr);
library(tidyr);
library(RColorBrewer);
library(dendextend);
library(gplots);
library(cluster);
```

```{r}
kuang <- read.csv("./datasets/kuang-2014-microarray-expression.csv") ##local- needs to be changed 
time <- read.csv("./datasets/kuang-2014-microarry-expression-timepts.csv")

head(kuang[0:5])  ##preview data sets
tail(kuang[0:5]) 
```

```{r}
rnaseq <- read.csv("./datasets/kuang-2014-rnaseq-expression.csv") #local needs to be changed
rnatime <- read.csv("./datasets/kuang-2014-rnaseq-expression-timepts.csv")
```

```{r}
dim(kuang)
```

```{r}
# first remember the names
n <- kuang$Gene.name
# transpose all but the first column (Gene.name)
kuang_t <- as.data.frame(t(kuang[,-1]))
colnames(kuang_t) <- n

dim(kuang_t)
dim(time)
kuang_t.time <- cbind(kuang_t, time = time$minute)
head(kuang_t.time[6241:6242])
```

```{r, echo = TRUE}
colnames(kuang_t) %>% length()
unique.genes <- colnames(kuang_t) %>% unique()
length(unique.genes)

```
While looking at the data set we observed multiple genes that were present more than once.  Using the unique function we can see that there are 6241 genes in the data set, but only 6035 unique genes.  We are not sure for the reasoning behind this, our best guess is redundancy, but it is something to keep in mind.

## Time plots 

```{r, echo = TRUE, fig.align="center"}
#visual of first 10 genes
first.10.genes <- c("YAL001C", "YAL002W", "YAL003W", "YAL004W", "YAL005C", "YAL007C", "YAL008W", "YAL009W", "YAL010C", "YAL011W")

ggplot(kuang_t.time, aes(x = time, y = YAL001C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL001C")

ggplot(kuang_t.time, aes(x = time, y = YAL002W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL002W")

ggplot(kuang_t.time, aes(x = time, y = YAL003W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL003W")

ggplot(kuang_t.time, aes(x = time, y = YAL004W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL004W")

ggplot(kuang_t.time, aes(x = time, y = YAL005C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL005C")

ggplot(kuang_t.time, aes(x = time, y = YAL007C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL007C")

ggplot(kuang_t.time, aes(x = time, y = YAL008W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL008W")

ggplot(kuang_t.time, aes(x = time, y = YAL009W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL009W")

ggplot(kuang_t.time, aes(x = time, y = YAL010C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL010C")

ggplot(kuang_t.time, aes(x = time, y = YAL011W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL011W")
```

## Clustering 

**Complete method, with 1-Pearson correlation as the distance matrix**
 
First, we attempted hierarchical clustering using the complete method with 1-Pearson's correlation as the distance metric. The first thing we did was create the correlation matrix and then distance matrix of the transpose of the original data (with genes and the columns). We are clustered the transpose of the original matrix because we wanted to cluster genes based on their relative correlation over time. All of the clustering techniques we know cluster the columns, not rows, of the dataset. 

```{r}
## These take a lot of time so they are moved into their own code segment
kuang_t.cor <- kuang_t %>% cor(use="pairwise.complete.obs")  ##generate correlation measurements
kuang.dist <- as.dist(1 - kuang_t.cor)  ##use correlation to find distance measurements 
```

```{r}
kuang.tree <- hclust(kuang.dist, method="complete")
kuang.dend <- as.dendrogram(kuang.tree) ## created & priviewed dendrogram object
```

To get an idea of the data we plotted the dendrogram.  From the dendrogram it is clear that there are significant clusters in the dataset. 

```{r}
plot(kuang.dend, leaflab = "none")  ## removing labels 
```

To further get a feel for the dataset we plotted the dendrogram at several different cluster values (2, 4, and 8)- color coded to best analyze the data.

```{r}
clusters_2 <- cutree(kuang.dend, k=2)
plot(color_branches(kuang.dend, k=2),leaflab="none")
table(clusters_2)
```

```{r}
clusters_4 <- cutree(kuang.dend, k=4)
plot(color_branches(kuang.dend, k=4),leaflab="none")
table(clusters_4)
```

```{r}
clusters_8 <- cutree(kuang.dend, k=8)
plot(color_branches(kuang.dend, k=8),leaflab="none")
table(clusters_8)
```

```{r}
sub.trees <- cut(kuang.dend, h = 1.75)  ##Height retrieved from 4 cluster Figure above
cluster1.final <- sub.trees$lower[[1]]
cluster2.final <- sub.trees$lower[[2]]
cluster3.final <- sub.trees$lower[[3]]
cluster4.final <- sub.trees$lower[[4]]
cluster5.final <- sub.trees$lower[[5]]
cluster6.final <- sub.trees$lower[[6]]
cluster7.final <- sub.trees$lower[[7]]
cluster8.final <- sub.trees$lower[[8]]
nleaves(cluster1.final)
nleaves(cluster2.final)
nleaves(cluster3.final)
nleaves(cluster4.final)
nleaves(cluster5.final)
nleaves(cluster6.final)
nleaves(cluster7.final)
nleaves(cluster8.final)
```


After reading our paper, we saw that they found three major clusters.  We next did hierarchical clustering using k = 3 to see if our clusters align with the clusters found in the paper.

```{r}
clusters_3 <- cutree(kuang.dend, k=3)
plot(color_branches(kuang.dend, k=3),leaflab="none", main = "Original Cluster with k = 3")
table(clusters_3)
```
We then cut the tree to extract these three clusters.
```{r}
cut1 <- cut(kuang.dend, h = 1.9)
cut1$lower
```
Perfect.  The first cluster countains 2956 genes, the second contains 972 genes, and the third contains 2313 genes.

To get a better idea we performed hierarchical clustering, starting with the first cluster on the left (pink)
```{r}
plot(cut1$lower[[1]], leaflab="none", main = "Cluster 1 dendrogram")
```
We then tried plotting this at dendrogram at different k values to see what looks good.
```{r}
plot(color_branches(cut1$lower[[1]], k=3), leaflab="none", main = "Cluster 1 dendrogram with k = 3")
```
The first two clusters (pink and green) looked pretty compact to us, so we did not explore them anymore.  The third (blue) cluster appeared to have some significant subclusters in it, so we explored it more.

We saw that there appeared to be 4 main clusters.  Adding color allowed us to visualize if our observation held true.
```{r}
plot(color_branches(cluster1.cut1$lower[[3]], k=4), leaflab="none", main = "Cluster 1 subtree dendrogram with k = 4")
```
That looked good to us.  We then extracted these four clusters to see how large they were and what genes they contained.
```{r}
cluster1.cut1.cutAgain <- cut(cluster1.cut1$lower[[3]], h = 1.4)
cluster1.cut1.cutAgain$lower
```


Next we looked at the smallest cluster ([[2]]).  
```{r}
plot(color_branches(cut1$lower[[2]], k=2), leaflab="none", main = "Cluster 2 dendrogram with k = 2")
```
Looking at this we saw there were two main clusters.  We then cut this subtree again in order to extract these two subclusters.
```{r}
cut1.cut2 <- cut(cut1$lower[[2]], h = 1.6)
cut1.cut2$lower
```
The smaller cluster (green) had 208 genes in it.  This is a good size to work with, so we continued exploring this subcluster.  First we plotted it.
```{r}
plot(cut1.cut2$lower[[2]], leaflab = "none", main = "Subcluster of Cluster 2")
```
4 main clusters stuck out to us when looking at the above plot, so we plotted with color and k = 4.
```{r}
plot(color_branches(cut1.cut2$lower[[2]], k=4), leaflab="none", main = "Subcluster of Cluster 2 dendrogram with k = 4")
```
We then extracted these four clusters.
```{r}
cut1.cut2.cutTo4 <- cut(cut1.cut2$lower[[2]], h = 1.25)
#labels(cut1.cut2.cutTo4$lower[[1]])

```
The sizes looked reasonable to work with.

Lastly we analyzed the third ([[3]]) cluster (blue in the original dendrogram).

```{r}
plot(cut1$lower[[3]], leaflab="none", main = "Cluster 3 dendrogram")
```
On this dendrogram we tried plotting it with different k values to see which made the most sense.
```{r}
plot(color_branches(cut1$lower[[3]], k=5), leaflab="none", main = "Cluster 3 dendrogram with k = 5")
```
A k value of 5 looked good.  As we have before, we then extracted these clusters by cutting the dendrogram.
```{r}
cluster3.cut1 <- cut(cut1$lower[[3]], h = 1.6)
cluster3.cut1$lower
```

=======
##RNAseq clustering

```{r}
# first remember the names
nrna <- rnaseq$Gene.name

# transpose all but the first column (Gene.name)
rnaseq_t_zeros <- as.data.frame(t(rnaseq[,-1]))
colnames(rnaseq_t_zeros) <- nrna

#dim(rnaseq_t_zeros)
#dim(rnatime)
#tail(rnaseq_t_zeros[0:5])
rnaseq_t_zeros.time <- cbind(rnaseq_t_zeros, time = rnatime$hour)
#head(rnaseq_t_zeros.time[6241:6242])
```

## Time plots for rnaseq

```{r, echo = TRUE, fig.align="center"}
#visual of first 10 genes
first.10.genes <- c("YAL001C", "YAL002W", "YAL003W", "YAL004W", "YAL005C", "YAL007C", "YAL008W", "YAL009W", "YAL010C", "YAL011W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL001C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL001C")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL002W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL002W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL003W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL003W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL004W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL004W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL005C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL005C")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL007C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL007C")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL008W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL008W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL009W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL009W")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL010C, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL010C")

ggplot(rnaseq_t_zeros.time, aes(x = time, y = YAL011W, group = 1)) + geom_line() + labs(x = "Time (minutes)", y = "Expression of Gene", title = "YAL011W")
```

```{r}
dim(rnaseq_t_zeros)
head(rnaseq_t_zeros[,6240:6241])
```

Some of these genes have zero expression and need to be removed!

```{r}
head(rnaseq_t_zeros[, colSums(rnaseq_t_zeros == 0) > 0])
```

```{r}
rnaseq_t <- rnaseq_t_zeros[, colSums(rnaseq_t_zeros != 0) > 0]
head(rnaseq_t)
```

```{r}
rnaseq.cor <- cor(rnaseq_t, use = "pairwise.complete.obs") 
rnaseq.dist <- as.dist(1-rnaseq.cor)
```

```{r}
rnaseq.tree <- hclust(rnaseq.dist, method = "complete") 
rnaseq.dend <- as.dendrogram(rnaseq.tree)
```

```{r}
plot(rnaseq.dend, leaflab="none", main = "Gasch1k Dendrogram Complete Linkage")
```

